---
title: "Translation on Demand"
subtitle: "Progress Report"
author: "Elias Schwall"
format: revealjs
---

## RNA-binding proteins motif enrichment

1.  get the UTR sequences of ToD candidates using biomaRt

-   potential problem: retrievs UTR for each isoform -\> gene with more isoforms can potentially screw enrichment

2.  get the UTR of all genes (matched transcriptomic & proteomic data) for background using biomaRt
3.  run AME (Meme)
4.  retrieving the ToD candidates that carry the enriched motifs using `universalmotif::scan_sequences()`

## UTR motifs

1.  No meme databse available on UTR motifs

-   potential solution: curate custom motif list

## 

-   M307 (Nono): RNA splicing, DNA repair, transcription regulation, and immune response

```{r}
library(tidyverse)
ToD_genes_carrying_RBP_motifs <- here::here("results/meme_analysis", "ToD_genes_carrying_RBP_motifs.rds") %>% 
  readRDS()

library(gt)
ToD_genes_carrying_RBP_motifs$`1h_vs_0h`$utr_5 %>% 
  group_by(motif) %>% 
  gt() %>% 
  tab_header(title = "ToD genes carrying RBP motifs",
             subtitle = "in 5' UTRs of ToD Candidates 0h to 1h") %>% 
  opt_stylize(style = 3) %>% 
   tab_options(table.width = pct(100),  # Adjust table width
              table.font.size = 10,  # Adjust font size
              container.height = px(800), 
              container.overflow.y = "auto")
```

## 

::::: columns
::: {.column width="30%"}
-   M082 (Ybx2): The major constituent of messenger ribonucleoprotein particles (mRNPs) regulates germ cell mRNA stability and translation. It binds to the Y-box consensus promoter and full-length mRNA with high affinity. It also interacts with short RNA sequences (5'-UCCAUCA-3') with lower affinity. This protein is crucial for the cytoplasmic retention of maternal mRNAs and may mark Y-box promoter-transcribed mRNAs for cytoplasmic storage, linking transcription to mRNA storage and translational delay.
:::

::: {.column width="70%"}
```{r}
ToD_genes_carrying_RBP_motifs$`1h_vs_0h`$utr_3 %>% 
  group_by(motif) %>% 
  gt() %>% 
  tab_header(title = "ToD genes carrying RBP motifs",
             subtitle = "in 3' UTRs of ToD Candidates 0h to 1h") %>% 
  opt_stylize(style = 3) %>% 
   tab_options(table.width = pct(100),  # Adjust table width
              table.font.size = 10,  # Adjust font size
              container.height = px(800), 
              container.overflow.y = "auto")
```
:::
:::::

## 

```{r}
ToD_genes_carrying_RBP_motifs$`6h_vs_1h`$utr_5 %>% 
  group_by(motif) %>% 
  gt() %>% 
  tab_header(title = "ToD genes carrying RBP motifs",
             subtitle = "in 5' UTRs of ToD Candidates 1h to 6h") %>% 
  opt_stylize(style = 3) %>% 
   tab_options(table.width = pct(100),  # Adjust table width
              table.font.size = 10,  # Adjust font size
              container.height = px(800), 
              container.overflow.y = "auto")
```

## 

::::: columns
::: {.column width="30%"}
```{r}
 # Load the gt library
library(gt)

# Create a data frame with the gene information
gene_data <- data.frame(
  Gene = c("Aass", "Agap1", "Axin1", "Btbdb9", "Cacna1a", 
           "Cdc42bpa", "Cfap74", "Csrnp2", "Cttnbp2", "Dennd1b", 
           "Dnajc14", "Ebpl", "Fbxo9", "Flvcr2", "Gramd1b", 
           "Mlh3", "Otud5", "Pknox1", "Rassf1", "Raver2", 
           "Rgs19", "Rhob", "Rsrp1", "Slain1", "Stil", 
           "Taf11", "Toml2", "Ttlil5", "Zfp850", "Zik1"),
  Function = c("Involved in amino acid metabolism",
                           "Regulates actin cytoskeleton dynamics",
                           "Plays a role in Wnt signaling pathway",
                           "Involved in protein-protein interactions",
                           "Encodes a calcium channel, important for neurotransmission",
                           "Involved in cell signaling and cytoskeletal organization",
                           "Associated with cilia and flagella function",
                           "Involved in RNA binding and regulation",
                           "Plays a role in Wnt signaling and cell adhesion",
                           "Involved in endosomal trafficking",
                           "Heat shock protein involved in protein folding",
                           "Involved in mitochondrial function",
                           "Part of the ubiquitin-proteasome system",
                           "Involved in heme transport",
                           "Associated with lipid metabolism",
                           "Involved in DNA repair processes",
                           "Plays a role in ubiquitin signaling",
                           "Involved in transcriptional regulation",
                           "Associated with cell cycle regulation",
                           "Involved in RNA splicing and transport",
                           "Regulates G protein signaling",
                           "Involved in cytoskeletal dynamics",
                           "Plays a role in RNA splicing",
                           "Involved in cytoskeletal organization",
                           "Associated with cell cycle regulation",
                           "Involved in transcription regulation",
                           "Associated with mitochondrial function",
                           "Involved in RNA processing",
                           "Transcription factor involved in gene regulation",
                           "Involved in cellular signaling")
)

# Create the gt table
gene_data %>%
  gt()  %>% 
  tab_header(title = "ToDC carrying M082 (Ybx2) target motif") %>% 
  opt_stylize(style = 3) %>% 
   tab_options(table.width = pct(100),  # Adjust table width
              table.font.size = 10,  # Adjust font size
              container.height = px(800), 
              container.overflow.y = "auto")

```
:::

::: {.column width="70%"}
-   Cellular Functions: cellular processes such as signaling, transcription regulation, and cytoskeletal dynamics.

-   Metabolism: metabolic pathways, particularly related to amino acids and lipids

-   Development and Growth: developmental processes and cell cycle regulation.

-   Protein Interactions
:::
:::::

