---
title: "Translation on Demand"
subtitle: "Progress Report"
author: "Elias Schwall"
format: revealjs
---

## RNA-binding proteins motif enrichment
1. get the UTR sequences of ToD candidates using biomaRt 
  - potential problem: retrievs UTR for each isoform -> gene with more isoforms can potentially screw enrichment
2. get the UTR of all genes (matched transcriptomic & proteomic data) for background using biomaRt
3. run AME (Meme)
4. retrieving the ToD candidates that carry the enriched motifs using `universalmotif::scan_sequences()`



## UTR motifs
1. No meme databse available on UTR motifs 
- potential solution: curate custom motif list

## 
- M307 (Nono): RNA splicing, DNA repair, transcription regulation, and immune response
```{r}
library(tidyverse)
ToD_genes_carrying_RBP_motifs <- here::here("results/meme_analysis", "ToD_genes_carrying_RBP_motifs.rds") %>% 
  readRDS()

library(gt)
ToD_genes_carrying_RBP_motifs$`1h_vs_0h`$utr_5 %>% 
  group_by(motif) %>% 
  gt() %>% 
  tab_header(title = "ToD genes carrying RBP motifs",
             subtitle = "in 5' UTRs of ToD Candidates 0h to 1h") %>% 
  opt_stylize(style = 3) %>% 
   tab_options(table.width = pct(100),  # Adjust table width
              table.font.size = 10,  # Adjust font size
              container.height = px(800), 
              container.overflow.y = "auto")
```

## 
:::: {.columns}

::: {.column width="30%"}
- M082 (Ybx2): The major constituent of messenger ribonucleoprotein particles (mRNPs) regulates germ cell mRNA stability and translation. It binds to the Y-box consensus promoter and full-length mRNA with high affinity. It also interacts with short RNA sequences (5'-UCCAUCA-3') with lower affinity. This protein is crucial for the cytoplasmic retention of maternal mRNAs and may mark Y-box promoter-transcribed mRNAs for cytoplasmic storage, linking transcription to mRNA storage and translational delay.
:::

::: {.column width="70%"}
```{r}
ToD_genes_carrying_RBP_motifs$`1h_vs_0h`$utr_3 %>% 
  group_by(motif) %>% 
  gt() %>% 
  tab_header(title = "ToD genes carrying RBP motifs",
             subtitle = "in 3' UTRs of ToD Candidates 0h to 1h") %>% 
  opt_stylize(style = 3) %>% 
   tab_options(table.width = pct(100),  # Adjust table width
              table.font.size = 10,  # Adjust font size
              container.height = px(800), 
              container.overflow.y = "auto")
```
:::

::::

##

```{r}
ToD_genes_carrying_RBP_motifs$`6h_vs_1h`$utr_5 %>% 
  group_by(motif) %>% 
  gt() %>% 
  tab_header(title = "ToD genes carrying RBP motifs",
             subtitle = "in 5' UTRs of ToD Candidates 1h to 6h") %>% 
  opt_stylize(style = 3) %>% 
   tab_options(table.width = pct(100),  # Adjust table width
              table.font.size = 10,  # Adjust font size
              container.height = px(800), 
              container.overflow.y = "auto")
```
