---
title: "Translation on Demand"
format: dashboard
logo: "CECAD-Logo-24-white.png"
theme: minty
embed-resources: true
scrollable: true
nav-buttons:
      - icon: github
        href: https://github.com/eliasschwall/translation_on_demand
---

```{r set up}
#| label: set up chunk

library(tidyverse)
library(gt)
```

# Data Overview 

## Row 
```{r update valuebox}
#| content: valuebox
#| title: "Latest update"
list(
  icon = "clock",
  color = "secondary",
  value = Sys.Date()
)
```

::: {.card title="Data Source"}
The data used for this project is from Yang et al., 2019: [Multi-omic Profiling Reveals Dynamics of the Phased Progression of Pluripotency](https://www.sciencedirect.com/science/article/pii/S2405471219301152?via%3Dihub).

- [Table S2](https://ars.els-cdn.com/content/image/1-s2.0-S2405471219301152-mmc3.xlsx): Temporal Proteomic Data (0min, 30min, 1h, 6h, 12h, 24h, 36h, 48h, and 72h)
- [Table S3](https://ars.els-cdn.com/content/image/1-s2.0-S2405471219301152-mmc4.xlsx): Temporal Transcriptomic (RNA-Seq) Data (0min, 1h, 6h, 12h, 24h, 36h, 48h, and 72h)

For the purpuse of our project we realigned the count data for the transcriptomic data with the M11 mouse genome using STAR. 
:::

## Row

```{r displaying source data proteome}
#| title: Proteome Data Head

proteomic_yang_data <- here::here("data/processed/yang_paper", "yang_proteomics.rds") %>% readRDS()

proteomic_yang_data %>% 
  head() %>% 
  gt()
```

## Row

```{r displaying source data transcriptome}
#| title: Transcriptome Data Head

transcriptomic_yang_data <- here::here("data/raw/yang_paper", "transcriptomic_counts.csv") |> read.csv()

transcriptomic_yang_data %>% 
  head() %>% 
  gt()
```

# ToD Filtering 

# GO Enrichment 

# TF Analysis

# Motif Analysis 

## Row

```{r}
#| title: RNA Binding Protein Target Motifs Enriched in ToD Candidates

readRDS(here::here("results/meme_analysis", "RBP_motif_enrichmets_df.rds")) %>% 
  group_by(time_interval) %>% 
  gt() %>% 
  tab_header(title = "RNA Binding Protein Enriched Motifs",
             subtitle = "in 5' and 3' UTRs of ToD Candidates") %>% 
  tab_spanner(label = "Meme Ame Statistics",
              columns = 6:17) 
```

## Row {.tabset title="ToD candidates carrying RBP motifs"} 
```{r}
#| title: 0h to 1h 5' UTRs

ToD_genes_carrying_RBP_motifs <- here::here("results/meme_analysis", "ToD_genes_carrying_RBP_motifs.rds") %>% 
  readRDS()

library(gt)

# Assuming ToD_genes_carrying_RBP_motifs is defined and contains your data
ToD_genes_carrying_RBP_motifs$`1h_vs_0h`$utr_5 %>%
  group_by(motif) %>%
  gt() %>%
  tab_header(
    title = "ToD genes carrying RBP motifs",
    subtitle = "in 5' UTRs of ToD Candidates 0h to 1h"
  ) %>%
  tab_options(
    table.width = pct(100),  # Adjust table width
    table.font.size = 10,  # Adjust font size
    container.height = px(800),
    container.overflow.y = "auto"
  )
```

```{r}
#| title: 0h to 1h 3' UTRs

ToD_genes_carrying_RBP_motifs$`1h_vs_0h`$utr_3 %>%
  group_by(motif) %>%
  gt() %>%
  tab_header(
    title = "ToD genes carrying RBP motifs",
    subtitle = "in 3' UTRs of ToD Candidates 0h to 1h"
  ) %>%
  tab_options(
    table.width = pct(100),  # Adjust table width
    table.font.size = 10,  # Adjust font size
    container.height = px(800),
    container.overflow.y = "auto"
  )
```

```{r}
#| title: 1h to 6h 5' UTRs

ToD_genes_carrying_RBP_motifs$`6h_vs_1h`$utr_5 %>%
  group_by(motif) %>%
  gt() %>%
  tab_header(
    title = "ToD genes carrying RBP motifs",
    subtitle = "in 5' UTRs of ToD Candidates 1h to 6h"
  ) %>%
  tab_options(
    table.width = pct(100),  # Adjust table width
    table.font.size = 10,  # Adjust font size
    container.height = px(800),
    container.overflow.y = "auto"
  )
```

```{r}
#| title: 6h to 12h 5' UTRs

ToD_genes_carrying_RBP_motifs$`12h_vs_6h`$utr_5 %>%
  group_by(motif) %>%
  gt() %>%
  tab_header(
    title = "ToD genes carrying RBP motifs",
    subtitle = "in 5' UTRs of ToD Candidates 6h to 12h"
  ) %>%
  tab_options(
    table.width = pct(100),  # Adjust table width
    table.font.size = 10,  # Adjust font size
    container.height = px(800),
    container.overflow.y = "auto"
  )
```

```{r}
#| title: 12h to 24h 5' UTRs

ToD_genes_carrying_RBP_motifs$`24h_vs_12h`$utr_5 %>%
  group_by(motif) %>%
  gt() %>%
  tab_header(
    title = "ToD genes carrying RBP motifs",
    subtitle = "in 5' UTRs of ToD Candidates 12h to 24h"
  ) %>%
  tab_options(
    table.width = pct(100),  # Adjust table width
    table.font.size = 10,  # Adjust font size
    container.height = px(800),
    container.overflow.y = "auto"
  )
```

```{r}
#| title: 24h to 36h 5' UTRs

ToD_genes_carrying_RBP_motifs$`36h_vs_24h`$utr_5 %>%
  group_by(motif) %>%
  gt() %>%
  tab_header(
    title = "ToD genes carrying RBP motifs",
    subtitle = "in 5' UTRs of ToD Candidates 24h to 36h"
  ) %>%
  tab_options(
    table.width = pct(100),  # Adjust table width
    table.font.size = 10,  # Adjust font size
    container.height = px(800),
    container.overflow.y = "auto"
  )
```

```{r}
#| title: 36h to 48h 5' UTRs

ToD_genes_carrying_RBP_motifs$`48h_vs_36h`$utr_5 %>%
  group_by(motif) %>%
  gt() %>%
  tab_header(
    title = "ToD genes carrying RBP motifs",
    subtitle = "in 5' UTRs of ToD Candidates 36h to 48h"
  ) %>%
  tab_options(
    table.width = pct(100),  # Adjust table width
    table.font.size = 10,  # Adjust font size
    container.height = px(800),
    container.overflow.y = "auto"
  )
```

```{r}
#| title: 36h to 48h 3' UTRs

ToD_genes_carrying_RBP_motifs$`48h_vs_36h`$utr_3 %>%
  group_by(motif) %>%
  gt() %>%
  tab_header(
    title = "ToD genes carrying RBP motifs",
    subtitle = "in 3' UTRs of ToD Candidates 36h to 48h"
  ) %>%
  tab_options(
    table.width = pct(100),  # Adjust table width
    table.font.size = 10,  # Adjust font size
    container.height = px(800),
    container.overflow.y = "auto"
  )
```

```{r}
#| title: 48h to 72h 5' UTRs

ToD_genes_carrying_RBP_motifs$`72h_vs_48h`$utr_5 %>%
  group_by(motif) %>%
  gt() %>%
  tab_header(
    title = "ToD genes carrying RBP motifs",
    subtitle = "in 5' UTRs of ToD Candidates 48h to 72h"
  ) %>%
  tab_options(
    table.width = pct(100),  # Adjust table width
    table.font.size = 10,  # Adjust font size
    container.height = px(800),
    container.overflow.y = "auto"
  )
```

```{r}
#| title: 48h to 72h 3' UTRs

ToD_genes_carrying_RBP_motifs$`72h_vs_48h`$utr_3 %>%
  group_by(motif) %>%
  gt() %>%
  tab_header(
    title = "ToD genes carrying RBP motifs",
    subtitle = "in 3' UTRs of ToD Candidates 48h to 72h"
  ) %>%
  tab_options(
    table.width = pct(100),  # Adjust table width
    table.font.size = 10,  # Adjust font size
    container.height = px(800),
    container.overflow.y = "auto"
  )
```