---
title: "ToD Analysis of the realigned Yang data (mm11)"
author: "Elias"
format: html
editor: visual
---

We want to do our previous ToD analysis we did with the mm9 aligned Yang data with the new aligned mm11. In the last script we checked if the to data sets are comparable and from now on we are using the mm11 for further analysis.

First we want to calculate the TPM values of our count data.

```{r}
#| label: importing data

transcriptomic_yang_data <- read.csv("../data/processed/mm11_transcirpt_data.csv", row.names = 1)
proteomic_yang_data <- read.csv("../data/processed/proteomic_data_cleaned.csv")
```

```{r}
#| label: adding ensembl ids to the proteomic data
mart <- biomaRt::useEnsembl(biomart = "genes", dataset = "mmusculus_gene_ensembl")

ensembl_ids_for_proteomic_data <- biomaRt::getBM(attributes = c("mgi_symbol", 
                                                                "ensembl_gene_id"),
                                                 filters = "mgi_symbol",
                                                 values = proteomic_yang_data$GeneID,
                                                 mart = mart)

colnames(ensembl_ids_for_proteomic_data) <- c("GeneID", "ensembl_gene_id")

proteomic_yang_data_ensembl <- merge(proteomic_yang_data, ensembl_ids_for_proteomic_data, by="GeneID") %>% 
  rowwise() %>%
  mutate(row_mean = mean(c_across(3:10), na.rm = TRUE)) %>%
  group_by(ensembl_gene_id) %>%
  slice_max(order_by = row_mean, n = 1) %>%
  ungroup() %>%
  dplyr::select(-row_mean) %>% 
  tibble::column_to_rownames("ensembl_gene_id") %>% 
  dplyr::select(-c('UniprotID', "GeneID")) %>% 
  setNames(.,c("h0", "h1", "h6", "h12", "h24", "h36", "h48", "h72"))

rm(proteomic_yang_data)
rm(ensembl_ids_for_proteomic_data)
```

Now we have to normalize the transcirptomic data

```{r}
#| label: normalization of transcriptomic data

# getting the exons lengths of the genes 
exon_leghts <- read.csv("../data/reference/mouse_gene_exon_lengths.csv")


# calculating RPK
transcipt_data_RPK <- merge(transcriptomic_yang_data %>% tibble::rownames_to_column("ensembl_gene_id"), exon_leghts, by = "ensembl_gene_id") %>% 
  mutate(Exonic_Length = Exonic_Length / 1000) %>% 
  tibble::column_to_rownames('ensembl_gene_id') %>% 
  mutate(
    h0 = h0 / Exonic_Length,
    h1 = h1 / Exonic_Length,
    h6 = h6 / Exonic_Length,
    h12 = h12 / Exonic_Length,
    h24 = h24 / Exonic_Length,
    h36 = h36 / Exonic_Length,
    h48 = h48 / Exonic_Length,
    h72 = h72 / Exonic_Length
  ) %>% 
  dplyr::select(-Exonic_Length)

# calculating the scaling factors
TPM_scaling_factors <- colSums(transcipt_data_RPK) / 1e6

# calculating the TPM
transcipt_data_TPM <- sweep(transcipt_data_RPK, 2, TPM_scaling_factors, "/")

# taking the log2 of TMP values 
transcipt_data_TPM_log2 <- log2(transcipt_data_TPM + 1)

write.csv(transcipt_data_RPK, file = "../data/mm11/processed/transcipt_data_RPK.csv")
write.csv(transcipt_data_TPM, file = "../data/mm11/processed/transcipt_data_TPM.csv")
write.csv(transcipt_data_TPM_log2, file = "../data/mm11/processed/transcipt_data_TPM_log2.csv")
```

Next we calculate the fold changes for the transcipt_data_TPM_log2 and the proteomic_yang_data_ensembl

```{r}
#| label: calculating fold changes 
source("../R/calculating_FC_between_time_points.R")
transcipt_data_TPM_log2_FC <- calculating_FC_between_time_points(transcipt_data_TPM_log2)
proteomic_data_FC <- calculating_FC_between_time_points(proteomic_yang_data_ensembl)

write.csv(transcipt_data_TPM_log2_FC, file = "../data/mm11/processed/transcipt_data_TPM_log2_FC.csv")
write.csv(proteomic_data_FC, file = "../data/mm11/processed/proteomic_data_FC.csv")
```

Lastly, we want to calculate the ratio between protein fold change and RNA fold change

```{r}
#| label: calculating protein-RNA-FC-ratio

# merging the two dfs together 
protein_RNA_FCs <- merge(transcipt_data_TPM_log2_FC %>% tibble::rownames_to_column("ensembl_gene_id"),
                         proteomic_data_FC          %>% tibble::rownames_to_column("ensembl_gene_id"),
                         by = "ensembl_gene_id") %>% 
  rename_with(~ gsub("\\.x$", "_protein", .x)) %>%
  rename_with(~ gsub("\\.y$", "_RNA", .x)) %>% 
  tibble::column_to_rownames("ensembl_gene_id")

# calculating the rations
ratios_FC_protein_RNA <- protein_RNA_FCs[,1:7] - protein_RNA_FCs[,8:14] 
colnames(ratios_FC_protein_RNA) <- paste0(colnames(ratios_FC_protein_RNA)[1:7], "_RNA_ratio") 

write.csv(ratios_FC_protein_RNA, file = "../data/mm11/processed/ratios_FC_protein_RNA.csv")
```

The last thing to do, is to filter for the ToD candidates

```{r}
#| label: saving background genes 

# for enchrichment analysis we safe the background genes 
background_genes <- ratios_FC_protein_RNA %>% rownames() 

# we add the gene symbols for easy annotation later 
background_genes <- biomaRt::getBM(attributes = c("mgi_symbol", 
                                                  "ensembl_gene_id"),
                                   filters = "ensembl_gene_id",
                                   values = background_genes,
                                   mart = mart)

write.csv(background_genes, file = "../data/mm11/processed/background_genes.csv")



```

```{r}
#| label: filtering for ToD candidates 
candidate_threshold <- 1

ratios_FC_protein_RNA$GeneID <- rownames(ratios_FC_protein_RNA)


FC_0m_to_1h_protein_RNA_ratio_gene_list <- ratios_FC_protein_RNA$GeneID[ratios_FC_protein_RNA$FC_h0_to_h1_protein_RNA_ratio > candidate_threshold] %>% na.omit()

FC_1h_to_6h_protein_RNA_ratio_gene_list <- ratios_FC_protein_RNA$GeneID[ratios_FC_protein_RNA$FC_h1_to_h6_protein_RNA_ratio > candidate_threshold] %>% na.omit()

FC_6h_to_12h_protein_RNA_ratio_gene_list <- ratios_FC_protein_RNA$GeneID[ratios_FC_protein_RNA$FC_h6_to_h12_protein_RNA_ratio > candidate_threshold] %>% na.omit()

FC_12h_to_24h_protein_RNA_ratio_gene_list <- ratios_FC_protein_RNA$GeneID[ratios_FC_protein_RNA$FC_h12_to_h24_protein_RNA_ratio > candidate_threshold] %>% na.omit()

FC_24h_to_36h_protein_RNA_ratio_gene_list <- ratios_FC_protein_RNA$GeneID[ratios_FC_protein_RNA$FC_h24_to_h36_protein_RNA_ratio > candidate_threshold] %>% na.omit()

FC_36h_to_48h_protein_RNA_ratio_gene_list <- ratios_FC_protein_RNA$GeneID[ratios_FC_protein_RNA$FC_h36_to_h48_protein_RNA_ratio > candidate_threshold] %>% na.omit()

FC_48h_to_72h_protein_RNA_ratio_gene_list <- ratios_FC_protein_RNA$GeneID[ratios_FC_protein_RNA$FC_h48_to_h72_protein_RNA_ratio > candidate_threshold] %>% na.omit()

translation_on_demand_candidates <- list(
  TOD_candidates_0m_to_1h = FC_0m_to_1h_protein_RNA_ratio_gene_list,
  TOD_candidates_1h_to_6h = FC_1h_to_6h_protein_RNA_ratio_gene_list,
  TOD_candidates_6h_to_12h = FC_6h_to_12h_protein_RNA_ratio_gene_list,
  TOD_candidates_12h_to_24h = FC_12h_to_24h_protein_RNA_ratio_gene_list,
  TOD_candidates_24h_to_36h = FC_24h_to_36h_protein_RNA_ratio_gene_list,
  TOD_candidates_36h_to_48h = FC_36h_to_48h_protein_RNA_ratio_gene_list,
  TOD_candidates_48h_to_72h = FC_48h_to_72h_protein_RNA_ratio_gene_list
)

jsonlite::write_json(translation_on_demand_candidates, path = "../results/mm11_remapped/translation_on_demand_candidates.json")
```

Let's check if we have ToD candidates that are both in the mm9 and mm11 mapped data 
```{r}
#| label: comparing old and new ToD candidates 

old_tod_candidates <- jsonlite::read_json("../results/mm9_old_data/translation_on_demand_candidates_1_threshold.json", simplifyVector = T)

new_tod_candidates <- jsonlite::read_json("../results/mm11_remapped/translation_on_demand_candidates.json", simplifyVector = T)

background_genes$mgi_symbol[background_genes$ensembl_gene_id %in% new_tod_candidates$TOD_candidates_0m_to_1h] %in% old_tod_candidates$TOD_candidates_0m_to_1h %>% table()

background_genes$mgi_symbol[background_genes$ensembl_gene_id %in% new_tod_candidates$TOD_candidates_1h_to_6h] %in% old_tod_candidates$TOD_candidates_1h_to_6h %>% table()

background_genes$mgi_symbol[background_genes$ensembl_gene_id %in% new_tod_candidates$TOD_candidates_6h_to_12h] %in% old_tod_candidates$TOD_candidates_6h_to_12h %>% table()

background_genes$mgi_symbol[background_genes$ensembl_gene_id %in% new_tod_candidates$TOD_candidates_12h_to_24h] %in% old_tod_candidates$TOD_candidates_12h_to_24h %>% table()

background_genes$mgi_symbol[background_genes$ensembl_gene_id %in% new_tod_candidates$TOD_candidates_24h_to_36h] %in% old_tod_candidates$TOD_candidates_24h_to_36h %>% table()

background_genes$mgi_symbol[background_genes$ensembl_gene_id %in% new_tod_candidates$TOD_candidates_36h_to_48h] %in% old_tod_candidates$TOD_candidates_36h_to_48h %>% table()

background_genes$mgi_symbol[background_genes$ensembl_gene_id %in% new_tod_candidates$TOD_candidates_48h_to_72h] %in% old_tod_candidates$TOD_candidates_48h_to_72h %>% table()
```

